---
title: "User defined bins for time to change token aesthetics"
author: "Dominic Rowney"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette 
vignette: >
  %\VignetteIndexEntry{User defined bins for time to change token aesthetics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

It is possible to change the aesthetics of tokens based on the timestamp of the animation

For example, the number of days a 'patient' has been in the system

```{r echo = TRUE, message = FALSE}
# Libraries ---------------------------------------------------------------
library(dplyr)            ##pipes
library(tidyr)            ##tidy data, partcularly the crossing() function
library(lubridate)        ##date time manipulation
library(bupaR)            ##buisness process analytics
library(processanimateR)  ##animates process

# Create performance time flags ------------------------------------------------
my_flags <- data.frame(value = c(0,2,4,8,16)) %>% 
            mutate(day = days(value)) #convert numeric value into days

```

This will change the colour of the token at 0, 2, 4, 8, and 16 days.

The `crossing()` function joins the cases of 'patients' to 'my_flags' and creates all possible combinations.

```{r echo = TRUE, message = FALSE}
# Create timestamps of flags ----------------------------------------------

my_timeflags <- patients %>% 
                cases %>%
                crossing(my_flags) %>% ##similar to a SQL outer join
                mutate(time = start_timestamp + day) %>% 
                filter(time <= complete_timestamp) %>% 
                select("case" = patient,time,value) ##must be case, time, value
```


The data for the `token_scale()` function must have the column headings 'case, time, value'.

Without the `domain = my_flags$value` arguement the flags follow alphabetic order (e.g. 
 0, 16, 2, 4, 8) rather than the numeric order we wants. See [d3-legend](https://github.com/susielu/d3-legend) for further information.

```{r echo = TRUE, message = FALSE}
# Animate process ---------------------------------------------------------

patients %>%
  animate_process(mode ="absolute",
                  jitter=10,
                  legend = "color", 
                  mapping = token_aes(
                    color = token_scale(my_timeflags
                                        , scale = "ordinal"
                                        , domain = my_flags$value
                                        , range = rev(RColorBrewer::brewer.pal(5,"Spectral"))
                    )))

```

The colors can be modified through the `range` argument. In this case the scale is reversed with `rev()` to go from blue to red. See `RColorBrewer::brewer.pal.info` for all options:

*Acknowledgement*

Thanks to [Dominic Rowney](https://github.com/DomRowney/) for this nice example of advanced processanimateR usage. The original example code can be found [here](https://github.com/DomRowney/ProcessAnimateR_vignette).
